---
title: "Week 2 Assignment, DATA-607"
author: "Fernando Figueres"
output: html_notebook
---

**Assignment: Upload movie review data into a MySQL database and then load the data into an R data frame.**

###Import Libraries
```{r Import Libraries}
library(DBI)
library(RMySQL)
```


###Connect to Cloud Database
```{r Connect to Database}
mdb <- DBI::dbConnect(RMySQL::MySQL(),
  dbname='movie_lens',
  host = "35.243.235.220",
  user = "root",
  password = rstudioapi::askForPassword("Database password")
)

#code borrowed from https://dbplyr.tidyverse.org/articles/dbplyr.html

#The data is based on the "Movielens 100k" dataset found at http://grouplens.org/datasets/movielens/100k/
```


###Define SQL query to extract required data
```{r}
sql <- "SELECT `items`.`title` AS `title`, `items`.`date` AS `release date`, `ratingsdata`.`reviewdatetime` AS `review date & time`, `ratingsdata`.`rating` AS `rating`, `users`.`gender` AS `gender`, `users`.`age` AS `age`, `users`.`occupation` AS `occupation`, `users`.`zip` AS `zip` FROM ((`ratingsdata` JOIN `items` ON ((`items`.`itemid` = `ratingsdata`.`itemid`))) JOIN `users` ON ((`ratingsdata`.`userid` = `users`.`userid`))) ORDER BY items.date DESC"
```


###Data is retreived and loaded into a data frame
```{r}
mdata <- dbGetQuery(mdb, sql)
```

```{r}
head(mdata)
```

###Sample graph
```{r}
hist(mdata$rating,main ="Rating Frequency")

```

###Bibliography

F. Maxwell Harper and Joseph A. Konstan. 2015. The MovieLens Datasets:
History and Context. ACM Transactions on Interactive Intelligent
Systems (TiiS) 5, 4, Article 19 (December 2015), 19 pages.
DOI=http://dx.doi.org/10.1145/2827872

